---
title: "606 HW Week 1"
author: "Kyle Gilde"
date: "February 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 606 HW Week 1

Practice: 1.7 (available in R using the data(iris) command), 1.9, 1.23, 1.33, 1.55, 1.69
Graded: 1.8, 1.10, 1.28, 1.36, 1.48, 1.50, 1.56, 1.70
For 1.48, the following R code will create a vector scores that can be used to answer the question:

scores <- c(57, 66, 69, 71, 72, 73, 74, 77, 78, 78, 79, 79, 81, 81, 82, 83, 83, 88, 89, 94)


1.1 Migraine and acupuncture. A migraine is a particularly painful type of headache, which
patients sometimes wish to treat with acupuncture. To determine whether acupuncture relieves
migraine pain, researchers conducted a randomized controlled study where 89 females diagnosed
with migraine headaches were randomly assigned to one of two groups: treatment or control.
43 patients in the treatment group received acupuncture that is specifically designed to treat
migraines. 46 patients in the control group received placebo acupuncture (needle insertion at nonacupoint
locations). 24 hours after patients received acupuncture, they were asked if they were
pain free. Results are summarized in the contingency table below.51
Pain free
Yes No Total
Treatment 10 33 43
Group
Control 2 44 46
Total 12 77 89
identified on the antero-internal part of the antitragus, the
anterior part of the lobe and the upper auricular concha, on
the same side of pain. The majority of these points were
effective very rapidly (within 1 min), while the remaining
points produced a slower antalgic response, between 2 and
5 min. The insertion of a semi-permanent needle in these
zones allowed stable control of the migraine pain, which
occurred within 30 min and still persisted 24 h later.
Since the most active site in controlling migraine pain
was the antero-internal part of the antitragus, the aim of
this study was to verify the therapeutic value of this elective
area (appropriate point) and to compare it with an area
of the ear (representing the sciatic nerve) which is probably
inappropriate in terms of giving a therapeutic effect on
migraine attacks, since it has no somatotopic correlation
with head pain.
Materials and methods
The study enrolled 94 females, diagnosed as migraine
without aura following the International Classification of
Headache Disorders [5], who were subsequently examined
at the Women’s Headache Centre, Department of Gynaecology
and Obstetrics of Turin University. They were all
included in the study during a migraine attack provided that
it started no more than 4 h previously. According to a
predetermined computer-made randomization list, the eligible
patients were randomly and blindly assigned to the
following two groups: group A (n = 46) (average age
35.93 years, range 15–60), group B (n = 48) (average age
33.2 years, range 16–58).
Before enrollment, each patient was asked to give an
informed consent to participation in the study.
Migraine intensity was measured by means a VAS
before applying NCT (T0).
In group A, a specific algometer exerting a maximum
pressure of 250 g (SEDATELEC, France) was chosen to
identify the tender points with Pain–Pressure Test (PPT).
Every tender point located within the identified area by the
pilot study (Fig. 1, area M) was tested with NCT for 10 s
starting from the auricle, that was ipsilateral, to the side of
prevalent cephalic pain. If the test was positive and the
reduction was at least 25% in respect to basis, a semipermanent
needle (ASP SEDATELEC, France) was
inserted after 1 min. On the contrary, if pain did not lessen
after 1 min, a further tender point was challenged in the
same area and so on. When patients became aware of an
initial decrease in the pain in all the zones of the head
affected, they were invited to use a specific diary card to
score the intensity of the pain with a VAS at the following
intervals: after 10 min (T1), after 30 min (T2), after
60 min (T3), after 120 min (T4), and after 24 h (T5).
In group B, the lower branch of the anthelix was
repeatedly tested with the algometer for about 30 s to
ensure it was not sensitive. On both the French and Chinese
auricular maps, this area corresponds to the representation
of the sciatic nerve (Fig. 1, area S) and is specifically used
to treat sciatic pain. Four needles were inserted in this area,
two for each ear.
In all patients, the ear acupuncture was always performed
by an experienced acupuncturist. The analysis of
the diaries collecting VAS data was conducted by an
impartial operator who did not know the group each patient
was in.
The average values of VAS in group A and B were
calculated at the different times of the study, and a statistical
evaluation of the differences between the values
obtained in T0, T1, T2, T3 and T4 in the two groups
studied was performed using an analysis of variance
(ANOVA) for repeated measures followed by multiple
t test of Bonferroni to identify the source of variance.
Moreover, to evaluate the difference between group B
and group A, a t test for unpaired data was always performed
for each level of the variable ‘‘time’’. In the case of
proportions, a Chi square test was applied. All analyses
were performed using the Statistical Package for the Social
Sciences (SPSS) software program. All values given in the
following text are reported as arithmetic mean (±SEM).
Results
Only 89 patients out of the entire group of 94 (43 in group
A, 46 in group B) completed the experiment. Four patients
withdrew from the study, because they experienced an
unbearable exacerbation of pain in the period preceding the
last control at 24 h (two from group A and two from group
B) and were excluded from the statistical analysis since
they requested the removal of the needles. One patient
from group A did not give her consent to the implant of the
semi-permanent needles. In group A, the mean number of
Fig. 1 The appropriate area
(M) versus the inappropriate
area (S) used in the treatment
of migraine attacks
S174 Neurol Sci (2011) 32 (Suppl 1):S173–S175
123
Figure from the original
paper displaying the appropriate
area (M) versus
the inappropriate area (S)
used in the treatment of
migraine attacks.
(a) What percent of patients in the treatment group were pain free 24 hours after receiving
acupuncture? What percent in the control group?
(b) At first glance, does acupuncture appear to be an e↵ective treatment for migraines? Explain
your reasoning.
(c) Do the data provide convincing evidence that there is a real pain reduction for those patients
in the treatment group? Or do you think that the observed di↵erence might just be due to
chance?
1.2 Sinusitis and antibiotics. Researchers studying the e↵ect of antibiotic treatment for acute
sinusitis compared to symptomatic treatments randomly assigned 166 adults diagnosed with acute
sinusitis to one of two groups: treatment or control. Study participants received either a 10-day
course of amoxicillin (an antibiotic) or a placebo similar in appearance and taste. The placebo
consisted of symptomatic treatments such as acetaminophen, nasal decongestants, etc. At the end
of the 10-day period patients were asked if they experienced significant improvement in symptoms.
The distribution of responses is summarized below.52
Self-reported significant
improvement in symptoms
Yes No Total
Treatment 66 19 85
Group
Control 65 16 81
Total 131 35 166
(a) What percent of patients in the treatment group experienced a significant improvement in
symptoms? What percent in the control group?
(b) Based on your findings in part (a), which treatment appears to be more e↵ective for sinusitis?
(c) Do the data provide convincing evidence that there is a di↵erence in the improvement rates
of sinusitis symptoms? Or do you think that the observed di↵erence might just be due to
chance?
51G. Allais et al. “Ear acupuncture in the treatment of migraine attacks: a randomized trial on the
efficacy of appropriate versus inappropriate acupoints”. In: Neurological Sci. 32.1 (2011), pp. 173–175.
52J.M. Garbutt et al. “Amoxicillin for Acute Rhinosinusitis: A Randomized Controlled Trial”. In:
JAMA: The Journal of the American Medical Association 307.7 (2012), pp. 685–692.
56 CHAPTER 1. INTRODUCTION TO DATA
1.9.2 Data basics
1.3 Air pollution and birth outcomes, study components. Researchers collected data to
examine the relationship between air pollutants and preterm births in Southern California. During
the study air pollution levels were measured by air quality monitoring stations. Specifically, levels
of carbon monoxide were recorded in parts per million, nitrogen dioxide and ozone in parts per
hundred million, and coarse particulate matter (PM10) in μg/m3. Length of gestation data were
collected on 143,196 births between the years 1989 and 1993, and air pollution exposure during
gestation was calculated for each birth. The analysis suggested that increased ambient PM10 and,
to a lesser degree, CO concentrations may be associated with the occurrence of preterm births.53.
Identify
(a) the cases,
(b) the variables and their types, and
(c) the main research question
in this study.
1.4 Buteyko method, study components. The Buteyko method is a shallow breathing technique
developed by Konstantin Buteyko, a Russian doctor, in 1952. Anecdotal evidence suggests
that the Buteyko method can reduce asthma symptoms and improve quality of life. In a scientific
study to determine the e↵ectiveness of this method, researchers recruited 600 asthma patients aged
18-69 who relied on medication for asthma treatment. These patients were split into two research
groups: one practiced the Buteyko method and the other did not. Patients were scored on quality
of life, activity, asthma symptoms, and medication reduction on a scale from 0 to 10. On average,
the participants in the Buteyko group experienced a significant reduction in asthma symptoms
and an improvement in quality of life.54. Identify
(a) the cases,
(b) the variables and their types, and
(c) the main research question
in this study.
1.5 Cheaters, study components. Researchers studying the relationship between honesty,
age and self-control conducted an experiment on 160 children between the ages of 5 and 15.
Participants reported their age, sex, and whether they were an only child or not. The researchers
asked each child to toss a fair coin in private and to record the outcome (white or black) on a
paper sheet, and said they would only reward children who report white. Half the students were
explicitly told not to cheat and the others were not given any explicit instructions. In the no
instruction group probability of cheating was found to be uniform across groups based on child’s
characteristics. In the group that was explicitly told to not cheat, girls were less likely to cheat,
and while rate of cheating didn’t vary by age for boys, it decreased with age for girls.55 Identify
(a) the cases,
(b) the variables and their types, and
(c) the main research question
in this study.
53B. Ritz et al. “E↵ect of air pollution on preterm birth among children born in Southern California
between 1989 and 1993”. In: Epidemiology 11.5 (2000), pp. 502–511.
54J. McGowan. “Health Education: Does the Buteyko Institute Method make a di↵erence?” In: Thorax
58 (2003).
55Alessandro Bucciol and Marco Piovesan. “Luck or cheating? A field experiment on honesty with
children”. In: Journal of Economic Psychology 32.1 (2011), pp. 73–78.
1.9. EXERCISES 57
1.6 Stealers, study components. In a study of the relationship between socio-economic class
and unethical behavior, 129 University of California undergraduates at Berkeley were asked to
identify themselves as having low or high social-class by comparing themselves to others with
the most (least) money, most (least) education, and most (least) respected jobs. They were also
presented with a jar of individually wrapped candies and informed that the candies were for
children in a nearby laboratory, but that they could take some if they wanted. After completing
some unrelated tasks, participants reported the number of candies they had taken. It was found
that those who were identified as upper-class took more candy than others.56 Identify
(a) the cases,
(b) the variables and their types, and
(c) the main research question
in this study.
1.7 Fisher’s irises. Sir Ronald Aylmer Fisher was an English statistician, evolutionary biologist,
and geneticist who worked on a data set that contained sepal length and width, and petal length
and width from three species of iris flowers (setosa, versicolor and virginica). There were 50 flowers
from each species in the data set.57
(a) How many cases were included in the data?
(b) How many numerical variables are included
in the data? Indicate what they are, and if
they are continuous or discrete.
(c) How many categorical variables are included
in the data, and what are they? List
the corresponding levels (categories).
Photo by Ryan Claussen
(http://flic.kr/p/6QTcuX)
CC BY-SA 2.0 license
1.8 Smoking habits of UK residents. A survey was conducted to study the smoking habits
of UK residents. Below is a data matrix displaying a portion of the data collected in this survey.
Note that “£” stands for British Pounds Sterling, “cig” stands for cigarettes, and “N/A” refers to
a missing component of the data.58
sex age marital grossIncome smoke amtWeekends amtWeekdays
1 Female 42 Single Under£2,600 Yes 12 cig/day 12 cig/day
2 Male 44 Single £10,400 to £15,600 No N/A N/A
3 Male 53 Married Above£36,400 Yes 6 cig/day 6 cig/day
...
...
...
...
...
...
...
...
1691 Male 40 Single £2,600 to £5,200 Yes 8 cig/day 8 cig/day
(a) What does each row of the data matrix represent?
(b) How many participants were included in the survey?
(c) Indicate whether each variable in the study is numerical or categorical. If numerical, identify
as continuous or discrete. If categorical, indicate if the variable is ordinal.
56P.K. Pi↵ et al. “Higher social class predicts increased unethical behavior”. In: Proceedings of the
National Academy of Sciences (2012).
57R.A Fisher. “The Use of Multiple Measurements in Taxonomic Problems”. In: Annals of Eugenics 7
(1936), pp. 179–188.
58National STEM Centre, Large Datasets from stats4schools.
58 CHAPTER 1. INTRODUCTION TO DATA
1.9.3 Overview of data collection principles
1.9 Air pollution and birth outcomes, scope of inference. Exercise 1.3 introduces a study
where researchers collected data to examine the relationship between air pollutants and preterm
births in Southern California. During the study air pollution levels were measured by air quality
monitoring stations. Length of gestation data were collected on 143,196 births between the years
1989 and 1993, and air pollution exposure during gestation was calculated for each birth.
(a) Identify the population of interest and the sample in this study.
(b) Comment on whether or not the results of the study can be generalized to the population, and
if the findings of the study can be used to establish causal relationships.
1.10 Cheaters, scope of inference. Exercise 1.5 introduces a study where researchers studying
the relationship between honesty, age, and self-control conducted an experiment on 160 children
between the ages of 5 and 15. The researchers asked each child to toss a fair coin in private and to
record the outcome (white or black) on a paper sheet, and said they would only reward children
who report white. Half the students were explicitly told not to cheat and the others were not given
any explicit instructions. Di↵erences were observed in the cheating rates in the instruction and no
instruction groups, as well as some di↵erences across children’s characteristics within each group.
(a) Identify the population of interest and the sample in this study.
(b) Comment on whether or not the results of the study can be generalized to the population, and
if the findings of the study can be used to establish causal relationships.
1.11 Buteyko method, scope of inference. Exercise 1.4 introduces a study on using the
Buteyko shallow breathing technique to reduce asthma symptoms and improve quality of life. As
part of this study 600 asthma patients aged 18-69 who relied on medication for asthma treatment
were recruited and randomly assigned to two groups: one practiced the Buteyko method and the
other did not. Those in the Buteyko group experienced, on average, a significant reduction in
asthma symptoms and an improvement in quality of life.
(a) Identify the population of interest and the sample in this study.
(b) Comment on whether or not the results of the study can be generalized to the population, and
if the findings of the study can be used to establish causal relationships.
1.12 Stealers, scope of inference. Exercise 1.6 introduces a study on the relationship between
socio-economic class and unethical behavior. As part of this study 129 University of California
Berkeley undergraduates were asked to identify themselves as having low or high social-class by
comparing themselves to others with the most (least) money, most (least) education, and most
(least) respected jobs. They were also presented with a jar of individually wrapped candies and
informed that the candies were for children in a nearby laboratory, but that they could take some if
they wanted. After completing some unrelated tasks, participants reported the number of candies
they had taken. It was found that those who were identified as upper-class took more candy than
others.
(a) Identify the population of interest and the sample in this study.
(b) Comment on whether or not the results of the study can be generalized to the population, and
if the findings of the study can be used to establish causal relationships.
1.9. EXERCISES 59
1.13 Relaxing after work. The 2010 General Social Survey asked the question, “After an
average work day, about how many hours do you have to relax or pursue activities that you
enjoy?” to a random sample of 1,155 Americans. The average relaxing time was found to be 1.65
hours. Determine which of the following is an observation, a variable, a sample statistic, or a
population parameter.
(a) An American in the sample.
(b) Number of hours spent relaxing after an average work day.
(c) 1.65.
(d) Average number of hours all Americans spend relaxing after an average work day.
1.14 Cats on YouTube. Suppose you want to estimate the percentage of videos on YouTube
that are cat videos. It is impossible for you to watch all videos on YouTube so you use a random
video picker to select 1000 videos for you. You find that 2% of these videos are cat videos.Determine
which of the following is an observation, a variable, a sample statistic, or a population parameter.
(a) Percentage of all videos on YouTube that are cat videos.
(b) 2%.
(c) A video in your sample.
(d) Whether or not a video is a cat video.
1.15 GPA and study hours. A survey was conducted on 193 Duke University undergraduates
who took an introductory statistics course in 2012. Among many other questions, this survey
asked them about their GPA, which can range between 0 and 4 points, and the number of hours
they spent studying per week. The scatterplot below displays the relationship between these two
variables.
(a) What is the explanatory variable and
what is the response variable?
(b) Describe the relationship between the
two variables. Make sure to discuss unusual
observations, if any.
(c) Is this an experiment or an observational
study?
(d) Can we conclude that studying longer
hours leads to higher GPAs?
Study hours/week
GPA
0 20 40 60
2.5
3.0
3.5
4.0
1.16 Income and education in US counties. The scatterplot below shows the relationship
between per capita income (in thousands of dollars) and percent of population with a bachelor’s
degree in 3,143 counties in the US in 2010.
(a) What are the explanatory and response
variables?
(b) Describe the relationship between the
two variables. Make sure to discuss unusual
observations, if any.
(c) Can we conclude that having a bachelor’s
degree increases one’s income?
Percent with Bachelor's degree
Per capita income (in $1,000)
10 30 50 70
20
40
60
60 CHAPTER 1. INTRODUCTION TO DATA
1.9.4 Observational studies and sampling strategies
1.17 Course satisfaction across sections. A large college class has 160 students. All 160
students attend the lectures together, but the students are divided into 4 groups, each of 40
students, for lab sections administered by di↵erent teaching assistants. The professor wants to
conduct a survey about how satisfied the students are with the course, and he believes that the
lab section a student is in might a↵ect the student’s overall satisfaction with the course.
(a) What type of study is this?
(b) Suggest a sampling strategy for carrying out this study.
1.18 Housing proposal across dorms. On a large college campus first-year students and
sophomores live in dorms located on the eastern part of the campus and juniors and seniors live in
dorms located on the western part of the campus. Suppose you want to collect student opinions
on a new housing structure the college administration is proposing and you want to make sure
your survey equally represents opinions from students from all years.
(a) What type of study is this?
(b) Suggest a sampling strategy for carrying out this study.
1.19 Internet use and life expectancy. The following scatterplot was created as part of
a study evaluating the relationship between estimated life expectancy at birth (as of 2014) and
percentage of internet users (as of 2009) in 208 countries for which such data were available.59
(a) Describe the relationship between life
expectancy and percentage of internet
users.
(b) What type of study is this?
(c) State a possible confounding variable
that might explain this relationship
and describe its potential e↵ect.
0 20 40 60 80 100
50
60
70
80
90
% Internet users
Life expectancy at birth
1.20 Stressed out, Part I. A study that surveyed a random sample of otherwise healthy high
school students found that they are more likely to get muscle cramps when they are stressed. The
study also noted that students drink more co↵ee and sleep less when they are stressed.
(a) What type of study is this?
(b) Can this study be used to conclude a causal relationship between increased stress and muscle
cramps?
(c) State possible confounding variables that might explain the observed relationship between
increased stress and muscle cramps.
1.21 Evaluate sampling methods. A university wants to determine what fraction of its
undergraduate student body support a new $25 annual fee to improve the student union. For each
proposed method below, indicate whether the method is reasonable or not.
(a) Survey a simple random sample of 500 students.
(b) Stratify students by their field of study, then sample 10% of students from each stratum.
(c) Cluster students by their ages (e.g. 18 years old in one cluster, 19 years old in one cluster,
etc.), then randomly sample three clusters and survey all students in those clusters.
59CIA Factbook, Country Comparisons, 2014.
1.9. EXERCISES 61
1.22 Random digit dialing. The Gallup Poll uses a procedure called random digit dialing,
which creates phone numbers based on a list of all area codes in America in conjunction with the
associated number of residential households in each area code. Give a possible reason the Gallup
Poll chooses to use random digit dialing instead of picking phone numbers from the phone book.
1.23 Haters are gonna hate, study confirms. A study published in the Journal of Personality
and Social Psychology asked a group of 200 randomly sampled men and women to evaluate how
they felt about various subjects, such as camping, health care, architecture, taxidermy, crossword
puzzles, and Japan in order to measure their dispositional attitude towards mostly independent
stimuli. Then, they presented the participants with information about a new product: a microwave
oven. This microwave oven does not exist, but the participants didn’t know this, and were given
three positive and three negative fake reviews. People who reacted positively to the subjects on
the dispositional attitude measurement also tended to react positively to the microwave oven, and
those who reacted negatively also tended to react negatively to it. Researchers concluded that
“some people tend to like things, whereas others tend to dislike things, and a more thorough
understanding of this tendency will lead to a more thorough understanding of the psychology of
attitudes.”60
(a) What are the cases?
(b) What is (are) the response variable(s) in this study?
(c) What is (are) the explanatory variable(s) in this study?
(d) Does the study employ random sampling?
(e) Is this an observational study or an experiment? Explain your reasoning.
(f) Can we establish a causal link between the explanatory and response variables?
(g) Can the results of the study be generalized to the population at large?
1.24 Family size. Suppose we want to estimate household size, where a “household” is defined
as people living together in the same dwelling, and sharing living accommodations. If we select
students at random at an elementary school and ask them what their family size is, will this be
a good measure of household size? Or will our average be biased? If so, will it overestimate or
underestimate the true value?
1.25 Flawed reasoning. Identify the flaw(s) in reasoning in the following scenarios. Explain
what the individuals in the study should have done di↵erently if they wanted to make such strong
conclusions.
(a) Students at an elementary school are given a questionnaire that they are asked to return after
their parents have completed it. One of the questions asked is, “Do you find that your work
schedule makes it difficult for you to spend time with your kids after school?” Of the parents
who replied, 85% said “no”. Based on these results, the school officials conclude that a great
majority of the parents have no difficulty spending time with their kids after school.
(b) A survey is conducted on a simple random sample of 1,000 women who recently gave birth,
asking them about whether or not they smoked during pregnancy. A follow-up survey asking
if the children have respiratory problems is conducted 3 years later, however, only 567 of these
women are reached at the same address. The researcher reports that these 567 women are
representative of all mothers.
(c) An orthopedist administers a questionnaire to 30 of his patients who do not have any joint
problems and finds that 20 of them regularly go running. He concludes that running decreases
the risk of joint problems.
60Justin Hepler and Dolores Albarrac´ın. “Attitudes without objects - Evidence for a dispositional
attitude, its measurement, and its consequences”. In: Journal of personality and social psychology 104.6
(2013), p. 1060.
62 CHAPTER 1. INTRODUCTION TO DATA
1.26 City council survey. A city council has requested a household survey be conducted in
a suburban area of their city. The area is broken into many distinct and unique neighborhoods,
some including large homes, some with only apartments, and others a diverse mixture of housing
structures. Identify the sampling methods described below, and comment on whether or not you
think they would be e↵ective in this setting.
(a) Randomly sample 50 households from the city.
(b) Divide the city into neighborhoods, and sample 20 households from each neighborhood.
(c) Divide the city into neighborhoods, randomly sample 10 neighborhoods, and sample all households
from those neighborhoods.
(d) Divide the city into neighborhoods, randomly sample 10 neighborhoods, and then randomly
sample 20 households from those neighborhoods.
(e) Sample the 200 households closest to the city council offices.
1.27 Sampling strategies. A statistics student who is curious about the relationship between
the amount of time students spend on social networking sites and their performance at school
decides to conduct a survey. Various research strategies for collecting data are described below.
In each, name the sampling method proposed and any bias you might expect.
(a) He randomly samples 40 students from the study’s population, gives them the survey, asks
them to fill it out and bring it back the next day.
(b) He gives out the survey only to his friends, making sure each one of them fills out the survey.
(c) He posts a link to an online survey on Facebook and asks his friends to fill out the survey.
(d) He randomly samples 5 classes and asks a random sample of students from those classes to fill
out the survey.
1.28 Reading the paper. Below are excerpts from two articles published in the NY Times:
(a) An article titled Risks: Smokers Found More Prone to Dementia states the following:61
“Researchers analyzed data from 23,123 health plan members who participated in a voluntary
exam and health behavior survey from 1978 to 1985, when they were 50-60 years old. 23 years
later, about 25% of the group had dementia, including 1,136 with Alzheimer’s disease and 416
with vascular dementia. After adjusting for other factors, the researchers concluded that pack-aday
smokers were 37% more likely than nonsmokers to develop dementia, and the risks went up
with increased smoking; 44% for one to two packs a day; and twice the risk for more than two
packs.”
Based on this study, can we conclude that smoking causes dementia later in life? Explain your
reasoning.
(b) Another article titled The School Bully Is Sleepy states the following:62
“The University of Michigan study, collected survey data from parents on each child’s sleep habits
and asked both parents and teachers to assess behavioral concerns. About a third of the students
studied were identified by parents or teachers as having problems with disruptive behavior or
bullying. The researchers found that children who had behavioral issues and those who were
identified as bullies were twice as likely to have shown symptoms of sleep disorders.”
A friend of yours who read the article says, “The study shows that sleep disorders lead to
bullying in school children.” Is this statement justified? If not, how best can you describe the
conclusion that can be drawn from this study?
61R.C. Rabin. “Risks: Smokers Found More Prone to Dementia”. In: New York Times (2010).
62T. Parker-Pope. “The School Bully Is Sleepy”. In: New York Times (2011).
1.9. EXERCISES 63
1.29 Shyness on Facebook. Given the anonymity a↵orded to individuals in online interactions,
researchers hypothesized that shy individuals might have more favorable attitudes toward
Facebook, and that shyness might be positively correlated with time spent on Facebook. They
also hypothesized that shy individuals might have fewer Facebook “friends” as they tend to have
fewer friends than non-shy individuals have in the o✏ine world. 103 undergraduate students at an
Ontario university were surveyed via online questionnaires. The study states “Participants were
recruited through the university’s psychology participation pool. After indicating an interest in
the study, participants were sent an e-mail containing the study’s URL.” Are the results of this
study generalizable to the population of all Facebook users?63
1.9.5 Experiments
1.30 Stressed out, Part II. In a study evaluating the relationship between stress and muscle
cramps, half the subjects are randomly assigned to be exposed to increased stress by being placed
into an elevator that falls rapidly and stops abruptly and the other half are left at no or baseline
stress.
(a) What type of study is this?
(b) Can this study be used to conclude a causal relationship between increased stress and muscle
cramps?
1.31 Light and exam performance. A study is designed to test the e↵ect of light level on
exam performance of students. The researcher believes that light levels might have di↵erent e↵ects
on males and females, so wants to make sure both are equally represented in each treatment. The
treatments are fluorescent overhead lighting, yellow overhead lighting, no overhead lighting (only
desk lamps).
(a) What is the response variable?
(b) What is the explanatory variable? What are its levels?
(c) What is the blocking variable? What are its levels?
1.32 Vitamin supplements. In order to assess the e↵ectiveness of taking large doses of vitamin
C in reducing the duration of the common cold, researchers recruited 400 healthy volunteers from
sta↵ and students at a university. A quarter of the patients were assigned a placebo, and the
rest were evenly divided between 1g Vitamin C, 3g Vitamin C, or 3g Vitamin C plus additives
to be taken at onset of a cold for the following two days. All tablets had identical appearance
and packaging. The nurses who handed the prescribed pills to the patients knew which patient
received which treatment, but the researchers assessing the patients when they were sick did not.
No significant di↵erences were observed in any measure of cold duration or severity between the
four medication groups, and the placebo group had the shortest duration of symptoms.64
(a) Was this an experiment or an observational study? Why?
(b) What are the explanatory and response variables in this study?
(c) Were the patients blinded to their treatment?
(d) Was this study double-blind?
(e) Participants are ultimately able to choose whether or not to use the pills prescribed to them.
We might expect that not all of them will adhere and take their pills. Does this introduce a
confounding variable to the study? Explain your reasoning.
63E.S. Orr et al. “The influence of shyness on the use of Facebook in an undergraduate sample”. In:
CyberPsychology & Behavior 12.3 (2009), pp. 337–340.
64C. Audera et al. “Mega-dose vitamin C in treatment of the common cold: a randomised controlled
trial”. In: Medical Journal of Australia 175.7 (2001), pp. 359–362.
64 CHAPTER 1. INTRODUCTION TO DATA
1.33 Light, noise, and exam performance. A study is designed to test the e↵ect of light
level and noise level on exam performance of students. The researcher believes that light and noise
levels might have di↵erent e↵ects on males and females, so wants to make sure both are equally
represented in each treatment. The light treatments considered are fluorescent overhead lighting,
yellow overhead lighting, no overhead lighting (only desk lamps). The noise treatments considered
are no noise, construction noise, and human chatter noise.
(a) What is the response variable?
(b) How many factors are considered in this study? Identify them, and describe their levels.
(c) What is the role of the sex variable in this study?
1.34 Music and learning. You would like to conduct an experiment in class to see if students
learn better if they study without any music, with music that has no lyrics (instrumental), or with
music that has lyrics. Briefly outline a design for this study.
1.35 Soda preference. You would like to conduct an experiment in class to see if your classmates
prefer the taste of regular Coke or Diet Coke. Briefly outline a design for this study.
1.36 Exercise and mental health. A researcher is interested in the e↵ects of exercise on
mental health and he proposes the following study: Use stratified random sampling to ensure representative
proportions of 18-30, 31-40 and 41- 55 year olds from the population. Next, randomly
assign half the subjects from each age group to exercise twice a week, and instruct the rest not
to exercise. Conduct a mental health exam at the beginning and at the end of the study, and
compare the results.
(a) What type of study is this?
(b) What are the treatment and control groups in this study?
(c) Does this study make use of blocking? If so, what is the blocking variable?
(d) Does this study make use of blinding?
(e) Comment on whether or not the results of the study can be used to establish a causal relationship
between exercise and mental health, and indicate whether or not the conclusions can
be generalized to the population at large.
(f) Suppose you are given the task of determining if this proposed study should get funding.
Would you have any reservations about the study proposal?
1.37 Chia seeds and weight loss. Chia Pets – those terra-cotta figurines that sprout fuzzy
green hair – made the chia plant a household name. But chia has gained an entirely new reputation
as a diet supplement. In one 2009 study, a team of researchers recruited 38 men and divided them
randomly into two groups: treatment or control. They also recruited 38 women, and they randomly
placed half of these participants into the treatment group and the other half into the control group.
One group was given 25 grams of chia seeds twice a day, and the other was given a placebo. The
subjects volunteered to be a part of the study. After 12 weeks, the scientists found no significant
di↵erence between the groups in appetite or weight loss.65
(a) What type of study is this?
(b) What are the experimental and control treatments in this study?
(c) Has blocking been used in this study? If so, what is the blocking variable?
(d) Has blinding been used in this study?
(e) Comment on whether or not we can make a causal statement, and indicate whether or not we
can generalize the conclusion to the population at large.
65D.C. Nieman et al. “Chia seed does not promote weight loss or alter disease risk factors in overweight
adults”. In: Nutrition Research 29.6 (2009), pp. 414–418.
1.9. EXERCISES 65
1.9.6 Examining numerical data
1.38 Mammal life spans. Data were collected on life spans (in years) and gestation lengths
(in days) for 62 mammals. A scatterplot of life span versus length of gestation is shown below.66
(a) What type of an association is apparent
between life span and length of gestation?
(b) What type of an association would you
expect to see if the axes of the plot were
reversed, i.e. if we plotted length of
gestation versus life span?
(c) Are life span and length of gestation
independent? Explain your reasoning.
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●●
●
●
●
●
●
●
●
●
●
Gestation (days)
Life Span (years)
0 200 400 600
0
25
50
75
100
1.39 Associations. Indicate which of the plots show a
(a) positive association
(b) negative association
(c) no association
Also determine if the positive and
negative associations are linear or
nonlinear. Each part may refer to
more than one plot.
●
●
●
●
●
●
●
●
●
●
●
●●
●
●
●
●●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●●
●
●
(1)
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●●
●
●
●●
●
●
●
●
●
●
●
●
●●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
(2)
●●
●
●
●
●●
●●
●
●
●
●
●
●
●
●
●
●
●
●
●●●
●●●●
●
●●●●
●●
●
●●
●
●
●●
●
●●
●●
●
●
●●●
●
●●
●
●
●
●
●●
●
●●
●●
●●●
●
●
●
●●
●
●
●●●
●●
●
●
●
●●●
●
●
●
●
●
●
●
●
●
●
●●
●●
(3)
●
●
●
●
●
●
●
●
●●
●●
●
●
●
●
●
●
●
●
●
●●
●
●
●
●●
●
●
●
●
●
●
●
●
●
●
●●
●
●
●
●
●
●
●
●
●
●
●●
●
●
●
●
●
●
●
●
●
●
●
●●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●
●●
●
●
●
●
●●
●
●
●
●●●
●
●
●
(4)
1.40 Office productivity. Office productivity is relatively low when the employees feel no stress
about their work or job security. However, high levels of stress can also lead to reduced employee
productivity. Sketch a plot to represent the relationship between stress and productivity.
1.41 Reproducing bacteria. Suppose that there is only sufficient space and nutrients to
support one million bacterial cells in a petri dish. You place a few bacterial cells in this petri dish,
allow them to reproduce freely, and record the number of bacterial cells in the dish over time.
Sketch a plot representing the relationship between number of bacterial cells and time.
1.42 Sleeping in college. A recent article in a college newspaper stated that college students
get an average of 5.5 hrs of sleep each night. A student who was skeptical about this value decided
to conduct a survey by randomly sampling 25 students. On average, the sampled students slept
6.25 hours per night. Identify which value represents the sample mean and which value represents
the claimed population mean.
66T. Allison and D.V. Cicchetti. “Sleep in mammals: ecological and constitutional correlates”. In:
Arch. Hydrobiol 75 (1975), p. 442.
66 CHAPTER 1. INTRODUCTION TO DATA
1.43 Parameters and statistics. Identify which value represents the sample mean and which
value represents the claimed population mean.
(a) American households spent an average of about $52 in 2007 on Halloween merchandise such as
costumes, decorations and candy. To see if this number had changed, researchers conducted
a new survey in 2008 before industry numbers were reported. The survey included 1,500
households and found that average Halloween spending was $58 per household.
(b) The average GPA of students in 2001 at a private university was 3.37. A survey on a sample of
203 students from this university yielded an average GPA of 3.59 in Spring semester of 2012.
1.44 Make-up exam. In a class of 25 students, 24 of them took an exam in class and 1 student
took a make-up exam the following day. The professor graded the first batch of 24 exams and
found an average score of 74 points with a standard deviation of 8.9 points. The student who took
the make-up the following day scored 64 points on the exam.
(a) Does the new student’s score increase or decrease the average score?
(b) What is the new average?
(c) Does the new student’s score increase or decrease the standard deviation of the scores?
1.45 Days o↵ at a mining plant. Workers at a particular mining site receive an average of 35
days paid vacation, which is lower than the national average. The manager of this plant is under
pressure from a local union to increase the amount of paid time o↵. However, he does not want
to give more days o↵ to the workers because that would be costly. Instead he decides he should
fire 10 employees in such a way as to raise the average number of days o↵ that are reported by his
employees. In order to achieve this goal, should he fire employees who have the most number of
days o↵, least number of days o↵, or those who have about the average number of days o↵?
1.46 Medians and IQRs. For each part, compare distributions (1) and (2) based on their
medians and IQRs. You do not need to calculate these statistics; simply state how the medians
and IQRs compare. Make sure to explain your reasoning.
(a) (1) 3, 5, 6, 7, 9
(2) 3, 5, 6, 7, 20
(b) (1) 3, 5, 6, 7, 9
(2) 3, 5, 7, 8, 9
(c) (1) 1, 2, 3, 4, 5
(2) 6, 7, 8, 9, 10
(d) (1) 0, 10, 50, 60, 100
(2) 0, 100, 500, 600, 1000
1.47 Means and SDs. For each part, compare distributions (1) and (2) based on their means
and standard deviations. You do not need to calculate these statistics; simply state how the means
and the standard deviations compare. Make sure to explain your reasoning. Hint: It may be useful
to sketch dot plots of the distributions.
(a) (1) 3, 5, 5, 5, 8, 11, 11, 11, 13
(2) 3, 5, 5, 5, 8, 11, 11, 11, 20
(b) (1) -20, 0, 0, 0, 15, 25, 30, 30
(2) -40, 0, 0, 0, 15, 25, 30, 30
(c) (1) 0, 2, 4, 6, 8, 10
(2) 20, 22, 24, 26, 28, 30
(d) (1) 100, 200, 300, 400, 500
(2) 0, 50, 300, 550, 600
1.48 Stats scores. Below are the final exam scores of twenty introductory statistics students.
57, 66, 69, 71, 72, 73, 74, 77, 78, 78, 79, 79, 81, 81, 82, 83, 83, 88, 89, 94
Create a box plot of the distribution of these scores. The five number summary provided below
may be useful.
Min Q1 Q2 (Median) Q3 Max
57 72.5 78.5 82.5 94
1.9. EXERCISES 67
1.49 Infant mortality. The infant mortality rate is defined as the number of infant deaths per
1,000 live births. This rate is often used as an indicator of the level of health in a country. The
relative frequency histogram below shows the distribution of estimated infant death rates for 224
countries for which such data were available in 2014.67
(a) Estimate Q1, the median, and Q3
from the histogram.
(b) Would you expect the mean of this
data set to be smaller or larger than
the median? Explain your reasoning.
Infant mortality
0 20 40 60 80 100 120
0
0.1
0.2
0.3
0.4
1.50 Mix-and-match. Describe the distribution in the histograms below and match them to
the box plots.
(a)
50 60 70
(b)
0 50 100
(c)
0 2 4 6
(1)
0
2
4
6
(2)
55
60
65
70
(3)
0
20
40
60
80
100
1.51 Air quality. Daily air quality is measured by the air quality index (AQI) reported by
the Environmental Protection Agency. This index reports the pollution level and what associated
health e↵ects might be a concern. The index is calculated for five major air pollutants
regulated by the Clean Air Act and takes values from 0 to 300, where a higher value indicates
lower air quality. AQI was reported for a sample of 91 days in 2011 in Durham, NC.
The relative frequency histogram below shows the distribution of the AQI values on these days.68
(a) Estimate the median AQI value of this sample.
(b) Would you expect the mean AQI value of this sample
to be higher or lower than the median? Explain
your reasoning.
(c) Estimate Q1, Q3, and IQR for the distribution.
(d) Would any of the days in this sample be considered
to have an unusually low or high AQI? Explain
your reasoning.
Daily AQI
10 20 30 40 50 60
0
0.05
0.1
0.15
0.2
67CIA Factbook, Country Comparisons, 2014.
68US Environmental Protection Agency, AirData, 2011.
68 CHAPTER 1. INTRODUCTION TO DATA
1.52 Median vs. mean. Estimate the median for the 400 observations shown in the histogram,
and note whether you expect the mean to be higher or lower than the median.
40 50 60 70 80 90 100
0
20
40
60
80
1.53 Histograms vs. box plots. Compare the two plots below. What characteristics of the
distribution are apparent in the histogram and not in the box plot? What characteristics are
apparent in the box plot but not in the histogram?
5 10 15 20 25
0
50
100
150
200
5
10
15
20
25
1.54 Marathon winners. The histogram and box plots below show the distribution of finishing
times for male and female winners of the New York Marathon between 1970 and 1999.
2.0 2.4 2.8 3.2
0
10
20
Marathon times
2.0
2.4
2.8
3.2
(a) What features of the distribution are apparent in the histogram and not the box plot? What
features are apparent in the box plot but not in the histogram?
(b) What may be the reason for the bimodal distribution? Explain.
(c) Compare the distribution of marathon times for men and women based on the box plot shown
below.
2.0 2.4 2.8 3.2
Women
Men
(d) The time series plot shown below is another way to look at these data. Describe what is visible
in this plot but not in the others.
● ● ● ●
●
● ● ●
● ●
●
● ● ● ● ● ●
● ● ●
●
●
●
●
●
●
● ● ● ●
Marathon times
1970 1975 1980 1985 1990 1995 2000
2.0
2.4
2.8
3.2
●
Women
Men
1.9. EXERCISES 69
1.55 Distributions and appropriate statistics, Part I. For each of the following, state
whether you expect the distribution to be symmetric, right skewed, or left skewed. Also specify
whether the mean or median would best represent a typical observation in the data, and whether
the variability of observations would be best represented using the standard deviation or IQR.
Explain your reasoning.
(a) Number of pets per household.
(b) Distance to work, i.e. number of miles between work and home.
(c) Heights of adult males.
1.56 Distributions and appropriate statistics, Part II . For each of the following, state
whether you expect the distribution to be symmetric, right skewed, or left skewed. Also specify
whether the mean or median would best represent a typical observation in the data, and whether
the variability of observations would be best represented using the standard deviation or IQR.
Explain your reasoning.
(a) Housing prices in a country where 25% of the houses cost below $350,000, 50% of the houses
cost below $450,000, 75% of the houses cost below $1,000,000 and there are a meaningful
number of houses that cost more than $6,000,000.
(b) Housing prices in a country where 25% of the houses cost below $300,000, 50% of the houses
cost below $600,000, 75% of the houses cost below $900,000 and very few houses that cost
more than $1,200,000.
(c) Number of alcoholic drinks consumed by college students in a given week. Assume that most of
these students don’t drink since they are under 21 years old, and only a few drink excessively.
(d) Annual salaries of the employees at a Fortune 500 company where only a few high level
executives earn much higher salaries than all the other employees.
1.57 TV watchers. Students in an AP Statistics class were asked how many hours of television
they watch per week (including online streaming). This sample yielded an average of 4.71 hours,
with a standard deviation of 4.18 hours. Is the distribution of number of hours students watch
television weekly symmetric? If not, what shape would you expect this distribution to have?
Explain your reasoning.
1.58 Exam scores. The average on a history exam (scored out of 100 points) was 85, with a
standard deviation of 15. Is the distribution of the scores on this exam symmetric? If not, what
shape would you expect this distribution to have? Explain your reasoning.
1.59 Facebook friends. Facebook data indicate that 50% of Facebook users have 100 or more
friends, and that the average friend count of users is 190. What do these findings suggest about
the shape of the distribution of number of friends of Facebook users?69
1.60 A new statistic. The statistic ¯x
median can be used as a measure of skewness. Suppose we
have a distribution where all observations are greater than 0, xi > 0. What is the expected shape
of the distribution under the following conditions? Explain your reasoning.
(a) ¯x
median = 1
(b) ¯x
median < 1
(c) ¯x
median > 1
69Lars Backstrom. “Anatomy of Facebook”. In: Facebook Data Teams Notes (2011).
70 CHAPTER 1. INTRODUCTION TO DATA
1.61 Income at the co↵ee shop. The first histogram below shows the distribution of the
yearly incomes of 40 patrons at a college co↵ee shop. Suppose two new people walk into the co↵ee
shop: one making $225,000 and the other $250,000. The second histogram shows the new income
distribution. Summary statistics are also provided.
(1)
60000 62500 65000 67500 70000
(2)
60000 110000 160000 210000 260000
(1) (2)
n 40 42
Min. 60,680 60,680
1st Qu. 63,620 63,710
Median 65,240 65,350
Mean 65,090 73,300
3rd Qu. 66,160 66,540
Max. 69,890 250,000
SD 2,122 37,321
(a) Would the mean or the median best represent what we might think of as a typical income
for the 42 patrons at this co↵ee shop? What does this say about the robustness of the two
measures?
(b) Would the standard deviation or the IQR best represent the amount of variability in the
incomes of the 42 patrons at this co↵ee shop? What does this say about the robustness of the
two measures?
1.62 Midrange. The midrange of a distribution is defined as the average of the maximum and
the minimum of that distribution. Is this statistic robust to outliers and extreme skew? Explain
your reasoning
1.63 Commute times. The US census collects data on time it takes Americans to commute
to work, among many other variables. The histogram below shows the distribution of average
commute times in 3,143 US counties in 2010. Also shown below is a spatial intensity map of the
same data.
Mean work travel (in min)
10 20 30 40
0
100
200
4
19
>33
(a) Describe the numerical distribution and comment on whether or not a log transformation may
be advisable for these data.
(b) Describe the spatial distribution of commuting times using the map below.
1.9. EXERCISES 71
1.64 Hispanic population. The US census collects data on race and ethnicity of Americans,
among many other variables. The histogram below shows the distribution of the percentage of the
population that is Hispanic in 3,143 counties in the US in 2010. Also shown is a histogram of logs
of these values.
Hispanic %
0 20 40 60 80 100
0
500
1000
1500
2000
log(% Hispanic)
−2 −1 0 1 2 3 4
0
50
100
150
200
250
0
20
>40
(a) Describe the numerical distribution and comment on why we might want to use log-transformed
values in analyzing or modeling these data.
(b) What features of the distribution of the Hispanic population in US counties are apparent in
the map but not in the histogram? What features are apparent in the histogram but not the
map?
(c) Is one visualization more appropriate or helpful than the other? Explain your reasoning.
1.9.7 Considering categorical data
1.65 Antibiotic use in children. The bar plot and the pie chart below show the distribution of
pre-existing medical conditions of children involved in a study on the optimal duration of antibiotic
use in treatment of tracheitis, which is an upper respiratory infection.
Gastrointestinal
Immunocompromised
Genetic/metabolic
Neuromuscular
Trauma
Respiratory
Cardiovascular
Prematurity
Relative frequency
0.00 0.10 0.20 0.30
Gastrointestinal
Immunocompromised
Genetic/metabolic
Neuromuscular
Trauma
Respiratory
Cardiovascular
Prematurity
(a) What features are apparent in the bar plot but not in the pie chart?
(b) What features are apparent in the pie chart but not in the bar plot?
(c) Which graph would you prefer to use for displaying these categorical data?
72 CHAPTER 1. INTRODUCTION TO DATA
1.66 Views on immigration. 910 randomly sampled registered voters from Tampa, FL were
asked if they thought workers who have illegally entered the US should be (i) allowed to keep their
jobs and apply for US citizenship, (ii) allowed to keep their jobs as temporary guest workers but
not allowed to apply for US citizenship, or (iii) lose their jobs and have to leave the country. The
results of the survey by political ideology are shown below.70
Political ideology
Conservative Moderate Liberal Total
(i) Apply for citizenship 57 120 101 278
(ii) Guest worker 121 113 28 262
Response
(iii) Leave the country 179 126 45 350
(iv) Not sure 15 4 1 20
Total 372 363 175 910
(a) What percent of these Tampa, FL voters identify themselves as conservatives?
(b) What percent of these Tampa, FL voters are in favor of the citizenship option?
(c) What percent of these Tampa, FL voters identify themselves as conservatives and are in favor
of the citizenship option?
(d) What percent of these Tampa, FL voters who identify themselves as conservatives are also in
favor of the citizenship option? What percent of moderates share this view? What percent of
liberals share this view?
(e) Do political ideology and views on immigration appear to be independent? Explain your
reasoning.
1.67 Views on the DREAM Act. A random sample of registered voters from Tampa, FL were
asked if they support the DREAM Act, a proposed law which would provide a path to citizenship
for people brought illegally to the US as children. The survey also collected information on the
political ideology of the respondents. Based on the mosaic plot shown below, do views on the
DREAM Act and political ideology appear to be independent? Explain your reasoning.71
Conservative Moderate Liberal
Support
Not support
Not sure
70SurveyUSA, News Poll #18927, data collected Jan 27-29, 2012.
71SurveyUSA, News Poll #18927, data collected Jan 27-29, 2012.
1.9. EXERCISES 73
1.68 Raise taxes. A random sample of registered voters nationally were asked whether they
think it’s better to raise taxes on the rich or raise taxes on the poor. The survey also collected
information on the political party affiliation of the respondents. Based on the mosaic plot shown
below, do views on raising taxes and political affiliation appear to be independent? Explain your
reasoning.72
Democrat Republican Indep / Other
Raise taxes on the rich
Raise taxes on the poor
Not sure
1.9.8 Case study: gender discrimination
1.69 Side e↵ects of Avandia. Rosiglitazone is the active ingredient in the controversial type 2
diabetes medicine Avandia and has been linked to an increased risk of serious cardiovascular problems
such as stroke, heart failure, and death. A common alternative treatment is pioglitazone, the
active ingredient in a diabetes medicine called Actos. In a nationwide retrospective observational
study of 227,571 Medicare beneficiaries aged 65 years or older, it was found that 2,593 of the 67,593
patients using rosiglitazone and 5,386 of the 159,978 using pioglitazone had serious cardiovascular
problems. These data are summarized in the contingency table below.73
Cardiovascular problems
Yes No Total
Treatment
Rosiglitazone 2,593 65,000 67,593
Pioglitazone 5,386 154,592 159,978
Total 7,979 219,592 227,571
(a) Determine if each of the following statements is true or false. If false, explain why. Be careful:
The reasoning may be wrong even if the statement’s conclusion is correct. In such cases, the
statement should be considered false.
i. Since more patients on pioglitazone had cardiovascular problems (5,386 vs. 2,593), we can
conclude that the rate of cardiovascular problems for those on a pioglitazone treatment is
higher.
ii. The data suggest that diabetic patients who are taking rosiglitazone are more likely to
have cardiovascular problems since the rate of incidence was (2,593 / 67,593 = 0.038)
3.8% for patients on this treatment, while it was only (5,386 / 159,978 = 0.034) 3.4% for
patients on pioglitazone.
iii. The fact that the rate of incidence is higher for the rosiglitazone group proves that rosiglitazone
causes serious cardiovascular problems.
iv. Based on the information provided so far, we cannot tell if the di↵erence between the
rates of incidences is due to a relationship between the two variables or due to chance.
(See the next page for additional parts to this question.)
72Public Policy Polling, Americans on College Degrees, Classic Literature, the Seasons, and More, data
collected Feb 20-22, 2015.
73D.J. Graham et al. “Risk of acute myocardial infarction, stroke, heart failure, and death in elderly
Medicare patients treated with rosiglitazone or pioglitazone”. In: JAMA 304.4 (2010), p. 411. issn:
0098-7484.
74 CHAPTER 1. INTRODUCTION TO DATA
(b) What proportion of all patients had cardiovascular problems?
(c) If the type of treatment and having cardiovascular problems were independent, about how
many patients in the rosiglitazone group would we expect to have had cardiovascular problems?
(d) We can investigate the relationship between outcome and treatment in this study using a
randomization technique. While in reality we would carry out the simulations required for
randomization using statistical software, suppose we actually simulate using index cards. In
order to simulate from the independence model, which states that the outcomes were independent
of the treatment, we write whether or not each patient had a cardiovascular problem
on cards, shu✏ed all the cards together, then deal them into two groups of size 67,593 and
159,978. We repeat this simulation 1,000 times and each time record the number of people in
the rosiglitazone group who had cardiovascular problems. Use the relative frequency histogram
of these counts to answer (i)-(iii).
i. What are the claims being tested?
ii. Compared to the number calculated in part
(b), which would provide more support for
the alternative hypothesis, more or fewer patients
with cardiovascular problems in the
rosiglitazone group?
iii. What do the simulation results suggest about
the relationship between taking rosiglitazone
and having cardiovascular problems in diabetic
patients? Simulated rosiglitazone cardiovascular events
2250 2350 2450
0
0.1
0.2
1.70 Heart transplants. The Stanford University Heart Transplant Study was conducted to
determine whether an experimental heart transplant program increased lifespan. Each patient
entering the program was designated an official heart transplant candidate, meaning that he was
gravely ill and would most likely benefit from a new heart. Some patients got a transplant and
some did not. The variable transplant indicates which group the patients were in; patients in the
treatment group got a transplant and those in the control group did not. Another variable called
survived was used to indicate whether or not the patient was alive at the end of the study. Of
the 34 patients in the control group, 30 died. Of the 69 people in the treatment group, 45 died.74
control treatment
alive
dead
Survival Time (days)
control treatment
0
500
1000
1500
(a) Based on the mosaic plot, is survival independent of whether or not the patient got a transplant?
Explain your reasoning.
(See the next page for additional parts to this question.)
74B. Turnbull et al. “Survivorship of Heart Transplant Data”. In: Journal of the American Statistical
Association 69 (1974), pp. 74–80.
1.9. EXERCISES 75
(b) What do the box plots below suggest about the efficacy (e↵ectiveness) of the heart transplant
treatment.
(c) What proportion of patients in the treatment group and what proportion of patients in the
control group died?
(d) One approach for investigating whether or not the treatment is e↵ective is to use a randomization
technique.
i. What are the claims being tested?
ii. The paragraph below describes the set up for such approach, if we were to do it without
using statistical software. Fill in the blanks with a number or phrase, whichever is
appropriate.
We write alive on cards representing patients who were alive at
the end of the study, and dead on cards representing patients
who were not. Then, we shu✏e these cards and split them into two groups:
one group of size representing treatment, and another group of
size representing control. We calculate the di↵erence between
the proportion of dead cards in the treatment and control groups (treatment -
control) and record this value. We repeat this 100 times to build a distribution
centered at . Lastly, we calculate the fraction of simulations where
the simulated di↵erences in proportions are . If this fraction is low,
we conclude that it is unlikely to have observed such an outcome by chance and
that the null hypothesis should be rejected in favor of the alternative.
iii. What do the simulation results shown below suggest about the e↵ectiveness of the transplant
program?



```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
