---
title: "D607 Week 2 - Regex Problems"
author: "Kyle Gilde"
date: "Feb. 12, 2017"
output: 
  html_document:
    toc: true
    #code_folding: hide
---
<style>
div.beige pre { background-color:beige; }
</style>
 <div class = "beige">

#The Task
Please deliver links to an R Markdown file (in GitHub and rpubs.com) with solutions to problems 3 and 4 from chapter 8 of Automated Data Collection in R.  Problem 9 is extra credit.  

#Load Packages
```{r setup, results='hide'} 
knitr::opts_chunk$set(#echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE,
                      tidy=TRUE,
                      #comment = "",
                      dev="png", 
                      dev.args=list(type="cairo"))

#https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf

load.packages <- c("RCurl", "knitr", "tesseract","stringr","htmlTable")


ipak <- function(pkg){
    #FUNCTION SOURCE: https://gist.github.com/stevenworthington/3178163
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}
ipak(load.packages)
```


###Just for fun, I used the optical character recognition pkg to get the text off of the textbook image. 
####I wanted to see if it could be done. It worked for more than half of the text. 
```{r ocr}
# text <- ocr("https://raw.githubusercontent.com/kylegilde/D607-Data-Acquistion/master/week3_problems.jpg")
# my.list <- unlist(strsplit(text, "\n"))
```



#Problem 3
a. Use the tools of this chapter to rearrange the Vector so that all elements conform to the standard First_name last_name.
```{r 3a}
# sub("fopy","Copy", my.list[1])
# cat(my.list[5], sub("Jv","_n", my.list[6]))

raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert" 

names <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
names_df <- data.frame(names, stringsAsFactors = FALSE)
names_df$first_half <- str_trim(str_replace(str_extract(names_df$names,"(, ).+"),",",""))
names_df$last_half <- str_trim(str_replace(str_extract(names_df$names,".+(, )"),",",""))
names_df$first_last_name <- ifelse(is.na(names_df$first_half),names_df$names, str_c(names_df$first_half, " ",names_df$last_half))
kable(names_df[,c(1,4)])
```

b. Construct a logical vector indicating whether a character has a title (i.e., Rev. and Dr.).
```{r 3b}
#str_replace(my.list[7],"Lew","i.e.")
regex1 <- "[[:alpha:]]{2,3}\\."

names_df$has_title <- str_detect(names_df$first_last_name, regex1)
kable(names_df[,c(4,5)])


```
c. Construct a logical vector indicating whether a character has a second name.
```{r 3c}
#my.list[8]
regex2 <- "^[[:alpha:]]\\.\\s"

names_df$has_2nd_name <- str_detect(names_df$first_last_name, regex2)
kable(names_df[,c(4,6)])
```
#Problem 4
Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression."                                                  
a. [0-9]+\\$  **look for numeric characters one or more times, followed by a dollar sign**
```{r 4a}
#my.list[9:15]
a <- "[0-9]+\\$"
a.ex <- "3425452345adsfas12351$asdfasdf"
str_extract(a.ex, a)
```

b. \\b[a-z]{1,4}\\b **look for strings composed of between 1 and 4 lowercase letters that also are word edges at both ends**       
```{r 4b}

b <- "\\b[a-z]{1,4}\\b"
b.ex <- " adsfas 54 cs asdff dae4 asds"
str_extract_all(b.ex, b)


```
c. .*?\\.txt$  
**look for strings that have any character zero or more times and that end with .txt**
```{r 4c}

c <- ".*?\\.txt$"
c.ex <- "34w54%&$YTAas54.txt"
str_extract_all(c.ex, c)
```
d. \\d{2}/\\d{2}/\\d{4}  **look for strings that have exactly 2 digits, followed by a forward slash, then exactly 2 digits and a forward slash, followed by exactly 4 digits**
```{r 4d}
d <- "\\d{2}/\\d{2}/\\d{4}"
d.ex <- "1/1/2011 12-05-2025sas22/43/520033551"
str_extract_all(d.ex, d)
```
e. <(.+?)>.+?</\\1> **look for HTML start tag w/o attributes, its content and its end tag**
```{r 4e}
e <- "<(.+?)>.+?</\\1>"
e.ex <- "wg45qa4d<title> 12-05-2025sas22/43/520033551</title>dfasdfadsf"
str_extract_all(e.ex, e)
```


#Problem 9
One following code hides a secret message, Crack it whh R and regular expressions. Hint: Some of the 
characters are more revealing than others
```{r 9}
jibberish <- getURL("http://www.r-datacollection.com/materials/regex/code_exercise.txt")
answer <- unlist(str_extract_all(jibberish, "[A-Z.]"))
cat(str_replace(answer,"\\."," "))
```
</div>
