---
title: "607 Data Aquisition - Week 1"
author: "Kyle Gilde"
date: "Feb. 5, 2017"
output: 
  html_document:
    toc: true
    #code_folding: hide
---
<style>
div.beige pre { background-color:beige; }
</style>
 <div class = "beige">

####Your task is to study the dataset and the associated description of the data (i.e. "data dictionary"). You may need to look around a bit, but it's there! 
+ You should take the data, and create a data frame with a subset of the columns in the dataset. 
+ You should include the column that indicates edible or poisonous and three or four other columns. 
+ You should also add meaningful column names and replace the abbreviations used in the data-for example, in the appropriate column, "e" might become "edible." 

####Your deliverable is the R code to perform these transformation tasks



#Setup
##[Mushroom Documentation](https://archive.ics.uci.edu/ml/datasets/Mushroom)
```{r setup, echo=F} 
knitr::opts_chunk$set(#echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE,
                      #tidy=TRUE,
                      comment = "",
                      dev="png", 
                      #fig.width = 6, 
                      #fig.height = 4, 
                      #dpi=200, 
                      dev.args=list(type="cairo"))
require(RCurl)
require(plyr)
```


#Load & Sample Data
```{r load data}
url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data"
mush <- getURL(url)
my.mush <- read.csv(text=mush, header = F, sep = ",", stringsAsFactors = F)
head(my.mush)
```


#Subset & transform data
```{r transform}
new.mush <- my.mush[, c(1, 5, 22, 23)]

names(new.mush) <- c("edibility", "has_bruises", "population", "habitat")
#alternative: rename()

new.mush$edibility[new.mush$edibility == "e"] <- "edible"
new.mush$edibility[new.mush$edibility == "p"] <- "poisonous"

new.mush$has_bruises[new.mush$has_bruises == "t"] <- "TRUE"
new.mush$has_bruises[new.mush$has_bruises == "f"] <- "FALSE"

new.mush$population[new.mush$population == "a"] <- "abundant"
new.mush$population[new.mush$population == "c"] <- "clustered"
new.mush$population[new.mush$population == "n"] <- "numerous"
new.mush$population[new.mush$population == "s"] <- "scattered"
new.mush$population[new.mush$population == "v"] <- "several"
new.mush$population[new.mush$population == "y"] <- "solitary"

new.mush$habitat[new.mush$habitat == "g"] <- "grasses"
new.mush$habitat[new.mush$habitat == "l"] <- "leaves"
new.mush$habitat[new.mush$habitat == "m"] <- "meadows"
new.mush$habitat[new.mush$habitat == "p"] <- "paths"
new.mush$habitat[new.mush$habitat == "u"] <- "urban"
new.mush$habitat[new.mush$habitat == "w"] <- "waste"
new.mush$habitat[new.mush$habitat == "d"] <- "woods"

```


#Sample Data & Convert back to factors
```{r sample}
new.mush[] <- lapply(new.mush, factor)
head(new.mush)
str(new.mush)

```

#How is the poisonous mushroom population distributed proportionally?
```{r }
p.mush <- subset(new.mush$population, new.mush$edibility == "poisonous")

round(t(t(prop.table(table(p.mush)))),2)

```


</div>