---
title: "D607 Project 2 - Tidy Data"
author: "Kyle Gilde"
date: "Mar. 4, 2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

#The Task
Choose any three of the "wide" datasets identified in the Week 6 Discussion items. (You may use your own dataset; please don't use my Sample Post dataset, since that was used in your Week 6 assignment!) 
 
For each of the three chosen datasets:
1.  Create a .CSV file (or optionally, a MySQL database!) that includes all of the information included in the dataset. You're encouraged to use a "wide" structure similar to how the information appears in the discussion item, so that you can practice tidying and transformations as described below.
2.  Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data. [Most of your grade will be based on this step!]
3.  Perform the analysis requested in the discussion item.

##Load Packages
```{r setup} 
knitr::opts_chunk$set(#echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE,
                      tidy=F,
                      #comment = "",
                      dev="png", 
                      dev.args=list(type="cairo"))
#https://cran.r-project.org/web/packages/prettydoc/vignettes/
#https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf

load.packages <- c("stringr","prettydoc", "tidyr", "dplyr", "knitr")

ipak <- function(pkg){
    #FUNCTION SOURCE: https://gist.github.com/stevenworthington/3178163
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}

ipak(load.packages)
```

##Load data, fill in missing values & take a look at it
```{r data}
my_data <- read.csv("https://raw.githubusercontent.com/kylegilde/D607-Data-Acquistion/master/data-sets/numbersense_flight_data.csv", stringsAsFactors = F)

#Fill in the airline values and insert underscore into "on time"
my_data[2,1] <- my_data[1,1]
my_data[5,1] <- my_data[4,1]
my_data[,2] <- sapply(my_data[,2], str_replace, " ", "_")

kable(my_data)
```

##Time to tidy this data up!
```{r tidy}
tidy_data <- my_data %>% 
  na.omit() %>% 
  rename(airline = X, arrival_type = X.1) %>% 
  gather("arrival_city", "n", 3:7) %>% 
  spread(arrival_type, "n") %>% 
  mutate(total_arrivals = delayed + on_time, on_time_rate = on_time / total_arrivals) %>% 
  arrange(desc(total_arrivals))

tidy_data[,2] <- sapply(tidy_data[,2], str_replace, "\\.", " ")
```

##Let's take a look at the results
```{r tidied}
glimpse(tidy_data)
kable(tidy_data)
```

#Now let's do some basic analysis with dply
##Which airline & city had the best on-time arrival rate?
```{r}
tidy_data %>% 
  filter(on_time_rate == max(on_time_rate)) %>% 
  kable()

```

##Which airline had the better on-time arrival rate?
```{r analysis, tidy=F}
tidy_data %>% 
  group_by(airline) %>% 
  summarise(airline_on_time_rate = sum(on_time) / sum(total_arrivals)) %>% 
  filter(airline_on_time_rate == max(airline_on_time_rate)) %>% 
  kable()
```

##Let's rank the cities by their overall on-time arrival rates.
```{r}
tidy_data %>% 
  group_by(arrival_city) %>% 
  summarise(city_on_time_rate = sum(on_time) / sum(total_arrivals)) %>% 
  mutate(on_time_ranking = min_rank(desc(city_on_time_rate))) %>% 
  arrange(on_time_ranking) %>% 
  kable()
```

