---
title: "D607 Project 1: Create Table Using Regex "
author: "Kyle Gilde"
date: "Feb. 19, 2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

<style>
.lightgreen table {
  background-color:#eff8e5;
.beige {
  background-color:beige;  
}
</style>
<div class = "beige">

##The Task
The chart above describes arrival delays for two airlines across five destinations. Your task is to:

(1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above.
You're encouraged to use a "wide" structure similar to how the information appears above, so
that you can practice tidying and transformations as described below.
(2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy
and transform your data.
(3) Perform analysis to compare the arrival delays for the two airlines.
(4) Your code should be in an R Markdown file, posted to rpubs.com, and should include narrative
descriptions of your data cleanup work, analysis, and conclusions. Please include in your
homework submission:
??? The URL to the .Rmd file in your GitHub repository. and
??? The URL for your rpubs.com web page.

##Load Packages
```{r setup, results='hide'} 
knitr::opts_chunk$set(#echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE,
                      tidy=TRUE,
                      #comment = "",
                      dev="png", 
                      dev.args=list(type="cairo"))

#https://cran.r-project.org/web/packages/prettydoc/vignettes/
#https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf

load.packages <- c("stringr","prettydoc", "tidyr", "dplyr")


ipak <- function(pkg){
    #FUNCTION SOURCE: https://gist.github.com/stevenworthington/3178163
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}
ipak(load.packages)
```

##Load Data & take a look at it
```{r data}
my.data <- read.csv("https://raw.githubusercontent.com/kylegilde/D607-Data-Acquistion/master/data-sets/numbersense_flight_data.csv", stringsAsFactors = F)

my.data

```

##Time to tidy this data up!
```{r tidy}
names(my.data)[1:2] <- c("airline", "arrival_type")
my.data[2,1] <- my.data[1,1]
my.data[5,1] <- my.data[4,1]
my.data[,2] <- sapply(my.data[,2], str_replace, " ", "_")

my.data
tidy.data <- my.data %>% 
  na.omit() %>% 
  gather("arrival_city", "n", 3:7) %>% 
  spread(arrival_type, "n") %>% 
  mutate(total_flights = delayed + on_time, on_time_rate = on_time / total_flights) %>% 
  arrange(desc(total_flights))
tidy.data[,2] <- sapply(tidy.data[,2], str_replace, "\\.", " ")

tidy.data
```
