---
title: "D607 Wk07 HW - Parsing HTML, XML & JSON"
author: "Kyle Gilde"
date: "Mar. 14, 2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

#The Task
Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more than one author. For each book, include the title, authors, and two or three other attributes that you find interesting.

Take the information that you've selected about these three books, and separately create three files which store the book's information in HTML (using an html table), XML, and JSON formats (e.g. "books.html", "books.xml", and "books.json"). To help you better understand the different file structures, I'd prefer that you create each of these files "by hand" unless you're already very comfortable with the file formats.

Write R code, using your packages of choice, to load the information from each of the three sources into separate R data frames. **Are the three data frames identical?**

Your deliverable is the three source files and the R code. If you can, package your assignment solution up into an .Rmd file and publish to rpubs.com. [This will also require finding a way to make your three text files accessible from the web].

#Load Packages
```{r setup} 
knitr::opts_chunk$set(warning=FALSE, 
                      message=FALSE,
                      tidy=F,
                      #comment = "",
                      dev="png", 
                      dev.args=list(type="cairo"))
#https://cran.r-project.org/web/packages/prettydoc/vignettes/
#https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf

#create vector with all needed packages
load_packages <- c("RCurl","prettydoc", "stringr", "dplyr", "knitr", "janitor", "XML", "tidyr", "RJSONIO")

#see if we need to install any of them
new.pkg <- load_packages[!(load_packages %in% installed.packages()[, "Package"])]
if (length(new.pkg)) install.packages(new.pkg, dependencies = TRUE, warn.conflicts = FALSE)

#require
t(t(sapply(load_packages, require, character.only = TRUE, quietly = TRUE,  warn.conflicts = FALSE)))
#CODE SOURCE DOCUMENTATION: https://gist.github.com/stevenworthington/3178163
```

#1. HTML Table Parsing
##Load data, look at the HTML table & then its data frame
```{r html}
url_html <- getURLContent("https://raw.githubusercontent.com/kylegilde/D607-Data-Acquistion/master/data-sets/books/books.html")

#my_html_df <- htmlParse(url_html)

writeLines(url_html)

my_html_df <- url_html %>% 
  readHTMLTable(header=TRUE, as.data.frame = TRUE) %>% 
  data.frame() %>% 
  clean_names() 

colnames(my_html_df) <- str_replace(colnames(my_html_df),"null_", "")

kable(my_html_df)
```

#2. JSON Parsing 
##Load data, look at the JSON & then its data frame
```{r json}

url_json <- getURLContent("https://raw.githubusercontent.com/kylegilde/D607-Data-Acquistion/master/data-sets/books/books.json")

print("Is my JSON valid?")
isValidJSON("https://raw.githubusercontent.com/kylegilde/D607-Data-Acquistion/master/data-sets/books/books.json")

writeLines(url_json)

my_json_df <- fromJSON(url_json)

my_json_df <- do.call("rbind", lapply(my_json_df$`favorite recent books`, data.frame, stringsAsFactors = F))

my_json_df <- my_json_df %>%  
  clean_names() %>% 
  arrange(book_title)

kable(my_json_df)


```

#3. XML Parsing 
##Load data, look at the XML & then its data frame
```{r xml}
url_XML <- getURLContent("https://raw.githubusercontent.com/kylegilde/D607-Data-Acquistion/master/data-sets/books/books2.xml")

writeLines(url_XML)

my_XML_df <- url_XML %>% 
  xmlParse() %>% 
  xmlToDataFrame(stringsAsFactors = FALSE)

kable(my_XML_df, caption = "This does not look the same as the HTML data frame.\\n For the book with more than one author, the function concatenated all of them into a single cell.\\n Let's do a little bit of surgery to get the same result.")


my_XML_df2 <- my_XML_df %>% 
  mutate(authors = paste(str_replace_all(authors, "([a-z])([A-Z])", "\\1,\\2"))) %>% 
  separate(authors, c(paste0("author_", 1:4)), sep = ",") %>% 
  gather(author_num, author, author_1:author_4, na.rm = T) %>% 
  select(book_title, author, everything(), -author_num) %>% 
  arrange(book_title) 
  
kable(my_XML_df2, caption = "And now it looks like the HTML data frame")

my_json_df == my_html_df
my_html_df == my_XML_df2
```



