---
title: "D607 Wk13 - Migrating Data to MongoDB"
author: "Kyle Gilde"
date: "April 29, 2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
#The Task
For this assignment, you should take information from a relational database and migrate it to a NoSQL database of your own choosing. 

For the NoSQL database, you may use MongoDB (which we introduced in week 7), Neo4j, or another NoSQL database of your choosing.

Your migration process needs to be reproducible.  R code is encouraged, but not required.  

You should also briefly describe the advantages and disadvantages of storing the data in a relational database vs. your NoSQL database.


```{r knitr_options, echo=FALSE}
knitr::opts_chunk$set(
                      warning = FALSE
                      ,error = FALSE
                      ,message = FALSE
                      ,tidy = TRUE
                      #,cache = TRUE
                      )
```


#Loaded Packages
```{r load_packages, echo = FALSE, comment = ""} 
#create vector with all needed packages
load_packages <- c("prettydoc", "devtools", "rmongodb", "RMySQL", "knitr", "jsonlite")

#see if we need to install any of them
install_load <- function(pkg){
  #CODE SOURCE: https://gist.github.com/stevenworthington/3178163
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
}

#excute function and display the loaded packages
data.frame(t(t(install_load(load_packages))), fix.empty.names = FALSE)


#install_github(repo = "mongosoup/rmongodb")
```

#Create the MySQL Database from CSVs
```{r rdbms}
movieDB <- dbConnect(MySQL(), user="root", password="kmgkmg", dbname="movie_ratings", host="localhost", port=3306)

my.files <- c(
  "friends"
  ,"movies"
  ,"movies_ratings_friends"
  ,"ratings")

root <- "https://raw.githubusercontent.com/kylegilde/D607-Data-Acquistion/master/data-sets/movie_ratings/"

#Load my files as DFs and then create tables
create_tables <- function(file_names, root, filetype, database){
  for (i in 1:length(file_names)) {
    my.url <- paste0(root, file_names[i], filetype)
    x <- read.csv(my.url)
    dbWriteTable(database, file_names[i], x, overwrite = T, row.names = F)
  }
}

create_tables(my.files, root, ".csv", movieDB)
# dbGetQuery(movieDB, paste("describe", my.files[1]))
# dbGetQuery(movieDB, paste("describe", my.files[2]))
# dbGetQuery(movieDB, paste("describe", my.files[3]))
# dbGetQuery(movieDB, paste("describe", my.files[4]))


all_tables <- dbGetQuery(movieDB, "show tables")
kable(all_tables, caption ="Tables")
```


#Create the MongoDB Database and Collections from the created MySQL Database
```{r mongo}
my.mongo <- mongo.create(host = "localhost") 

create_collections <- function(mysql_database, mongo_obj, mongo_db) {
  if (mongo.is.connected(mongo_obj)){
    #get all tables in the DB
    table_list <- dbGetQuery(mysql_database, "show tables")[, 1]
    select_star <- "select * from"
    for (i in 1:length(table_list)) {
      #create a DF from the table
      df <- dbGetQuery(movieDB, paste(select_star, table_list[i]))
      #create bson list
      bson_list <- lapply(split(df, 1:nrow(df)), function(x) mongo.bson.from.JSON(toJSON(x)))
      #insert bson list into collection
      collection <- paste(mongo_db, table_list[i], sep = ".") 
      mongo.insert.batch(mongo_obj, collection, bson_list)
      #print one record
      print(mongo.find.one(mongo_obj, collection))
    }
    #print(mongo.get.database.collections(my.mongo, mongo_db))
  } else {
      return("No connection")
    }
      
}


create_collections(movieDB, my.mongo, "movieDB")


# mongo.get.database.collections(my.mongo, db = "movieDB")
# 
# 
# mongo.get.databases(my.mongo)
# 
# 
# mongo.find.all(my.mongo, "movieBD.friends")


```

#The Advantages of Using MongoDB

According to [Tutorialspoint.com](https://www.tutorialspoint.com/mongodb/mongodb_advantages.htm), MongoDB is advantageous because

+ it is schema-less. A collection can hold more than one type of document.
+ it is very easy to scale.


```{r end, collapse=T}
dbDisconnect(movieDB)
mongo.drop.database(my.mongo, db = "movieDB")
mongo.destroy(my.mongo)
```

