---
title: "D607 Project 2 - Tidy Data"
author: "Kyle Gilde"
date: "Mar. 14, 2017"
output: 
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

#The Task
Pick three of your favorite books on one of your favorite subjects. At least one of the books should have more than one author. For each book, include the title, authors, and two or three other attributes that you find interesting.

Take the information that you've selected about these three books, and separately create three files which store the book's information in HTML (using an html table), XML, and JSON formats (e.g. "books.html", "books.xml", and "books.json"). To help you better understand the different file structures, I'd prefer that you create each of these files "by hand" unless you're already very comfortable with the file formats.

Write R code, using your packages of choice, to load the information from each of the three sources into separate R data frames. **Are the three data frames identical?**

Your deliverable is the three source files and the R code. If you can, package your assignment solution up into an .Rmd file and publish to rpubs.com. [This will also require finding a way to make your three text files accessible from the web].

#Load Packages
```{r setup} 
knitr::opts_chunk$set(#echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE,
                      tidy=F,
                      #comment = "",
                      dev="png", 
                      dev.args=list(type="cairo"))
#https://cran.r-project.org/web/packages/prettydoc/vignettes/
#https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf

#create vector with all needed packages
load_packages <- c("RCurl","prettydoc", "tidyr", "dplyr", "knitr", "janitor", "XML")

#see if we need to install 
new.pkg <- load_packages[!(load_packages %in% installed.packages()[, "Package"])]
if (length(new.pkg)) install.packages(new.pkg, dependencies = TRUE, warn.conflicts = FALSE)

#require
sapply(load_packages, require, character.only = TRUE, quietly = TRUE)
#CODE SOURCE DOCUMENTATION: https://gist.github.com/stevenworthington/3178163
```

#1. HTML Table 
##Load data & take a look the data frame
```{r load}
url_html <- getURLContent("https://raw.githubusercontent.com/kylegilde/D607-Data-Acquistion/master/data-sets/books/books.html")

#my_html_table <- htmlParse(url_html)

my_html_table <- url_html %>% 
  readHTMLTable(header=TRUE, as.data.frame = TRUE) %>% 
  data.frame() %>% 
  clean_names() 

colnames(my_html_table) <- str_replace(colnames(my_html_table),"null_", "")  

my_html_table
kable(my_html_table)
```

#2. XML Table 
##Load data & take a look the data frame
```{r tidy_data}

url_XML <- getURLContent("https://raw.githubusercontent.com/kylegilde/D607-Data-Acquistion/master/data-sets/favorite_books.XML")

#my_XML_table <- XMLParse(url_XML)
xml
my_XML_table <- data.frame(readXMLTable(url_XML))

kable(my_XML_table)

```

##Analysis that can be done:
1. Which issue attracted the most difference (in voting pattern) between members of congress according to affiliation?

*My interpretation of the question: On which issue was each party the most divided?*
```{r ds1q1}
party_agreement %>% 
  filter(party == "democrat") %>% 
  arrange(agreement_rate) %>% 
  head(1) %>% 
  kable()

party_agreement %>% 
  filter(party == "republican") %>% 
  arrange(agreement_rate) %>% 
  head(1) %>% 
  kable()
```


2. Which issue showed the most common ground (in voting pattern) between members of congress according to affiliation?

*My interpretation of the question: On which issue was each party the most unified?*
```{r ds1q2}
party_agreement %>% 
  filter(party == "democrat") %>% 
  arrange(desc(agreement_rate)) %>% 
  head(1) %>% 
  kable()

party_agreement %>% 
  filter(party == "republican") %>% 
  arrange(desc(agreement_rate)) %>% 
  head(1) %>% 
  kable()
```

