---
title: "D607 Project #1"
author: "Kyle Gilde"
date: "Feb. 12, 2017"
output: 
  html_document:
    toc: true
    #code_folding: hide
---

<style>
.lightgreen table {
  background-color:#eff8e5;
.beige {
  background-color:beige;  
}
</style>
<div class = "beige">

#The Task
In this project, you're given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:

Player's Name, Player's State, Total Number of Points, Player's Pre-Rating, and Average Pre Chess Rating of Opponents

For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605

#Load Packages
```{r setup, results='hide'} 
knitr::opts_chunk$set(#echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE,
                      tidy=TRUE,
                      #comment = "",
                      dev="png", 
                      dev.args=list(type="cairo"))

#https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf

load.packages <- c("RCurl", "knitr","stringr","htmlTable")


ipak <- function(pkg){
    #FUNCTION SOURCE: https://gist.github.com/stevenworthington/3178163
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}
ipak(load.packages)
```



```{r load.data}

```

</div>