---
title: "D607 Project #1"
author: "Kyle Gilde"
date: "Feb. 12, 2017"
output: 
  html_document:
    toc: true
    #code_folding: hide
---

<style>
.lightgreen table {
  background-color:#eff8e5;
.beige {
  background-color:beige;  
}
</style>
<div class = "beige">

#The Task
In this project, you're given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:

Player's Name, Player's State, Total Number of Points, Player's Pre-Rating, and Average Pre Chess Rating of Opponents

For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605

#Load Packages
```{r setup, results='hide'} 
knitr::opts_chunk$set(#echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE,
                      tidy=TRUE,
                      #comment = "",
                      dev="png", 
                      dev.args=list(type="cairo"))

#https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf

load.packages <- c("RCurl", "knitr","stringr","htmlTable")


ipak <- function(pkg){
    #FUNCTION SOURCE: https://gist.github.com/stevenworthington/3178163
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}
ipak(load.packages)
```



```{r load.data}

my.data <- getURLContent("https://raw.githubusercontent.com/kylegilde/D607-Data-Acquistion/master/data-sets/tournamentinfo.txt")

#extract all the attribute values needed
num <- str_trim(unlist(str_extract_all(my.data, " {3,4}\\d{1,2} ")))
name <- data.frame(str_trim(unlist(str_extract_all(my.data, " [[:alpha:] -]{2,30} {6}"))), stringsAsFactors = F)
name <- name[2:65,1]
state <- str_trim(unlist(str_extract_all(my.data, " {3}[[:alpha:]]{2} ")))
total_points <- as.numeric(str_trim(unlist(str_extract_all(my.data, "\\d\\.\\d"))))
pre_rating <- as.numeric(str_trim(unlist(str_extract_all(str_extract_all(my.data, ": {1,2}\\d{3,4}( {3}|P)"),"\\d+"))))


#to extract the opponent numbers, create a list seperated by newline characters
my.rows <- unlist(str_extract_all(my.data, ".+\\n"))
#subset to the rows that have names in them since these rows have the opponent numbers in them
my.rows <- str_subset(my.rows, " [[:alpha:] -]{2,30} {6}")
# extract only the digits of opponent numbers
my.rows <- str_extract_all(my.rows, "((W|L|D) {2,3}\\d{1,2})")
# just the digits
my.rows <- str_extract_all(my.rows, "\\d{1,2}")

opponents <- data.frame(sapply(my.rows, str_c, collapse=","), stringsAsFactors = F)
#remove superfluous row
opponents <- opponents[2:65,1]

#create my df so far
chess.df <- data.frame(num, name, state, total_points,pre_rating, opponents, stringsAsFactors = F)


opponent_mean <- function(opp.list, opp.scores){
  #function to calculate the opponent pre-rating means
  list_len <- length(opp.list)
  mean.vec <- rep(NA,list_len)
  for (i in 1:list_len)
    {
    opps <- c(str_split(opp.list,",")[[i]])
    opp.mean <- mean(subset(opp.scores[,2], opp.scores[,1] %in% opps))
    mean.vec[i] <- opp.mean
    }
  return(mean.vec)
}
chess.df$opponent_pre_rating_mean <- round(opponent_mean(chess.df[,6], chess.df[,c(1,5)]))

final_chess_df <- chess.df[,c(2:5,7)]
write.table(final_chess_df,file="kgilde_chess_df.csv",sep=",")
```

</div>