---
title: "D607 Week 2 - Movie Ratings DB"
author: "Kyle Gilde"
date: "Feb. 12, 2017"
output: 
  html_document:
    toc: true
    code_folding: hide
---
<style>
div.beige pre { background-color:beige; }
</style>
 <div class = "beige">

#The Task

Choose six recent popular movies.  Ask at least five people that you know (friends, family, classmates, imaginary friends) to rate each of these movie that they have seen on a scale of 1 to 5.  Take the results (observations) and store them in a SQL database.  Load the information into an R dataframe.

Your deliverables should include your SQL scripts and your R Markdown code, posted to GitHub.

#Load Packages
```{r setup} 
knitr::opts_chunk$set(#echo=FALSE, 
                      warning=FALSE, 
                      message=FALSE,
                      tidy=TRUE,
                      comment = "",
                      dev="png", 
                      dev.args=list(type="cairo"))

load.packages <- c(
  "RMySQL"
  #, "htmlTable"
  , "knitr"
)
ipak <- function(pkg){
    #Function source: https://gist.github.com/stevenworthington/3178163
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)) 
        install.packages(new.pkg, dependencies = TRUE)
    sapply(pkg, require, character.only = TRUE)
}
ipak(load.packages)
```


#Connect to the DB and list the tables.
```{r Connect}
my.db <- dbConnect(MySQL(), user="root", password="kmgkmg", dbname="movie_ratings", host="localhost", port=3306)

my.tables <- dbGetQuery(my.db, "show tables")
kable(my.tables, caption ="Tables")

# friends <- dbGetQuery(my.db, "select * from friends")
# movies <- dbGetQuery(my.db, "select * from movies")
# movies_ratings_friends <- dbGetQuery(my.db, "select * from movies_ratings_friends")
# ratings <- dbGetQuery(my.db, "select * from ratings")
```



#Join all tables into one aggregate table. View the 1st 10 rows.
```{r aggtable}
agg_table <- dbGetQuery(my.db, 
          "select m.movie
          , f.name
          , r.rating 
          , r.rating_score
          from movies_ratings_friends mrf
          join movies m on mrf.movie_id = m.id
          join friends f on mrf.friend_id = f.id
          join ratings r on mrf.rating_id = r.id"
           )

kable(head(agg_table, 10))
```

#The rating scores are 1 through 5, 1 being the lowest score.
```{r ratings}
my.ratings <- dbGetQuery(my.db,
                         "select rating
                          , rating_score
                          from ratings
                          order by 2")

kable(my.ratings)

```


#Which movie received the highest average rating from my friends?
```{r highest_rating}

highest_rating <- 
          dbGetQuery(my.db, 
          "select m.movie
          , avg(r.rating_score) avg_score
          from movies_ratings_friends mrf
          join movies m on mrf.movie_id = m.id
          join ratings r on mrf.rating_id = r.id
          group by 1
          order by 2 desc"
           )

kable(highest_rating, caption = "We have a tie between La La Land and Hell or High Water!")
```

#Which of my friends gave out the lowest ratings on average?

```{r most_critical_friend}
most_critical_friend <- dbGetQuery(my.db, 
          "select f.name
          , avg(r.rating_score) avg_rating
          from movies_ratings_friends mrf
          join friends f on mrf.friend_id = f.id
          join ratings r on mrf.rating_id = r.id
          group by 1
          order by 2"
          )

kable(most_critical_friend, caption = "Tomas gave out the lowest ratings!")
```




# Disconnect from DB
```{r disconnect}

dbDisconnect(my.db)

```



</div>
